Technique for \earlywarning\ detection of compact binary coalescence and its implication for multi-messenger astronomy

### 2011-06-03 method.tex
Fixed some notation in svd section
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@858 ccdd05d5-56c1-461a-bf98-85f6876523b4

As described previously, the template banks are, by design, highly correlated.
It is possible to greatly reduce the number of filters required to achieve a
particular minimum match by designing an appropriate set of orthonormal {\em
basis templates}.  A purely numerical technique based on the application of the
singular value decomposition (\SVD) to inspiral waveforms is demonstrated
in~\cite{Cannon:2010p10398}.  Using the results of ~\cite{Cannon:2010p10398} we
establish that the filters of the previous section can be approximated to high
accuracy by the expansion in the singular value basis
%
\begin{equation}
y_{ik}(\tau) \approx \sum_j^{\numsvdtmps} \sigma_{jk} v_{ijk} u_{j}(\tau)
\label{eq:svddecomp}
\end{equation}
%
where $\sigma_{jk}$ is the $j^{\mathrm{th}}$ singular value for the
$k^{\mathrm{th}}$ time slice, $v_{ijk}$ is an orthogonal matrix for the
$k^{\mathrm{th}}$ time slice  and $u_{jk}$ is a new orthogonal basis filter set
for the $k^{\mathrm{th}}$ time slice.  The authors of \cite{Cannon:2010p10398}
showed that to high reconstruction accuracy far fewer filters are needed than
were in the original template bank. We find that when combined with the
time-slice decomposition, the number of \SVD\ filters, \numsvdtmps\ is much smaller
than the original number of filters \numtmps.  We combine \eqref{eq:svddecomp}
with \eqref{eq:time-slices} to arrive at \eqref{eq:decomp}.  In the next
section we compute the expected computational cost scaling of this
decomposition and compare it with the brute-force implementation in
\eqref{eq:SNRTD} and higher latency \fft\ methods.

>>>>

As described previously, the template banks are, by design, highly correlated.
It is possible to greatly reduce the number of filters required to achieve a
particular minimum match by designing an appropriate set of orthonormal {\em
basis templates}.  A purely numerical technique based on the application of the
singular value decomposition (\SVD) to inspiral waveforms is demonstrated
in~\cite{Cannon:2010p10398}.  Using the results of ~\cite{Cannon:2010p10398} we
establish that the time sliced templates of the previous section can be
approximated to high accuracy by the expansion
%
\begin{equation}
h_i^s[k] = \sum_{\mathclap{l=0}}^{\mathclap{M-1}} v_{il}^s \sigma_l^s u_l^s[k] \approx \sum_{\mathclap{l=0}}^{\mathclap{L^s-1}} v_{il}^s \sigma_l^s u_l^s[k]
\label{eq:svddecomp}
\end{equation}
%
where $\sigma_l^s$ is the $l^{\mathrm{th}}$ singular value for the
$s^{\mathrm{th}}$ time slice, $v_{il}^s$ is an orthogonal matrix for the
$s^{\mathrm{th}}$ time slice, and $u_l^s[k]$ is a new set of orthogonal basis filters
for the $s^{\mathrm{th}}$ time slice.  The authors of \cite{Cannon:2010p10398}
showed that to high reconstruction accuracy far fewer filters are needed than
were in the original template bank. We find that when combined with the
time-slice decomposition, the number of \SVD\ filters, \numsvdtmps\ is much smaller
than the original number of filters \numtmps.  We combine \eqref{eq:svddecomp}
with \eqref{eq:time-slices} to arrive at \eqref{eq:decomp}.  In the next
section we compute the expected computational cost scaling of this
decomposition and compare it with the brute-force implementation in
\eqref{eq:SNRTD} and higher latency \fft\ methods.
